### 2016-10-14 robust-software.tex
Submission to arXiv

\section*{Notes}

@@

Typically, scientific software is developed and used by a single person,
usually a graduate student or postdoc, and produces the intended results in their
hands~\cite{prins2015}. 
But what happens
when someone else wants to use their software? Everyone with
a few years of experience feels a bit nervous when told to use
another person's code to analyze their data.
Often, that software will be undocumented and work in unexpected
ways (if it works at all).  It will often rely on nonexistent paths or resources,
be tuned for a single dataset,
or simply be an older version than was used in published papers.
The new user is then faced with two unpalatable options:
hack the existing code to make it work, or start over. Being unable to
replicate results is so common that one publication refers to it as "a rite of
passage"~\cite{baker2016}.

The root cause of this problem is that most research software
isn't \emph{robust}. The difference between running and being
robust is the difference between ``works for me on my machine'' and
``works for other people on a cluster I've never used''. The lack of
robustness also leads to many duplicated efforts,
slowing the pace of research~\cite{prabhu2011,lawlor2015}.
Bioinformatics software repositories \cite{ison2016,brazas2012} catalogue dozens to
hundreds of tools to perform the same task. In the Bioinformatics
Links Directory as of 2016, there are 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002,Seemann2013}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

This problem is not unique to bioinformatics, or even to computing~\cite{baker2016}. 
Best practices in software engineering specifically aim to increase software
robustness. However, most bioinformaticians come from related disciplines and
are trained on-the-job or otherwise informally~\cite{prins2015,atwood2015}. Even
the existing training programs and initiatives rarely have the time to cover software engineering
in-depth, especially since the field is so broad and rapidly
developing~\cite{atwood2015,lawlor2015}. Regardless of whether or not the skills
and knowledge are taught, developing robust software is not directly rewarded
in science and funding is difficult to come by~\cite{prins2015}. Some proposed
solutions to this problem, in addition to restructuring educational programs,
include hiring dedicated software engineers~\cite{lawlor2015,sanders2008},
partnering with private sector or grassroots organizations~\cite{prins2015,ison2016},
or using specific technical tools like containerization or cloud
computing~\cite{afgan2016,howe2012}. Each of these requires time and, in some
cases, institutional change.

You don't need to be a full-time programmer to write robust software. In fact,
some of the best, most reliable pieces of software in many scientific
communities are written by researchers~\cite{prabhu2011,sanders2008}. 
The most widely-used software packages tend to adopt strong software
engineering approaches, have high standards of reproducibility, good testing
practices, and inter-institutional collaborations that grow into strong user
bases. In the bioinformatics community, Bioconductor and Galaxy follow this
path~\cite{gentleman2004,afgan2016}.

In particular, robust software:

@@

These simple rules are also necessary steps toward creating a reusable
library that can be shared and reused through a site like CPAN or CRAN.  They are
generic and can be applied to all languages, libraries, packages,
documentation styles, and operating systems for both closed-source and
open-source software.  
Whether the aim is as simple as sharing the
code with collaborators or as complex as using the software in a
production analysis environment, increasing the robustness of your
software decreases headaches all around.

\section{Have a README that explains in a few lines what the software does and what its dependencies are.}

The README is the first stop for most new users.
At a minimum, it needs to get a new user started and point them towards more
help, if they need it. 
Numerous guidelines exist on how to write good
READMEs\cite{Johnson1997,gnustandards}. We summarize the crucial points here.

\textbf{Explain what the software does:} At the beginning of the README,
explain what the software does in one or two sentences.
There's nothing more frustrating
than spending the time to download and install some software only to
find out that it doesn't do what you thought it did.

@@

\textbf{Provide compile/installation instructions:} If the software needs to be compiled
or installed, list those instructions in the README. Also
mention if you recommend they use a pre-compiled binary instead
through a system such as pip, yum or apt.

\textbf{List input and output files}: All possible input and output files
should be listed in this section. Do the files conform to a particular
standard, an extension of an existing format, or is it your own
format? If using a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
explicitly, listing all the required fields and acceptable values.
(You get bonus points if you include a script to convert between
a standard format and your file format). If there is no rigorous format
(such as with log files), show an example file, or the first few lines,
and explain what the sections mean.

Input files and their formats are included in most documentation, but
intermediate, auxiliary and log files are often missing.
\emph{All} files should be listed in the README, even those considered self-explanatory. Log
files are often full of valuable information that can be
mined for the user's specific purpose. If your users might need to know,
``Does this program report the percentage of reads trimmed to remove
adapter sequences?'' they should be able to check the README and confidently
say, ``Yes, it is in the log file''.

\textbf{State attributions and licensing:} Attributions are how you credit
your main contributors; licenses are how you want others to use and
credit your software. READMEs are usually available \emph{before}
the software is installed (or even downloaded).
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
Attributions can also contain a list of ``expert'' users
that can be contacted if new users have problems with the software.

@@

Usage information provides the first line of help for both first-time and
experienced users of command-line applications.
Ideally, usage is a terse, informative command-line help message that
guides the user in the correct use of your software. Terseness is
important: usage that extends for multiple screens is a nuisance, especially when
printed to standard error instead of standard output (where it can
easily be paged).

Usage should provide all of the information necessary to run the
software. It is invoked either by running the software without
any arguments; running the software with incorrect arguments; or by
explicitly choosing a help or usage option.
More standard command-line behaviours are detailed in \cite{Seemann2013}.

@@

\textbf{Most commonly used arguments, a description of each, and the
default values}: Not all arguments need to appear in the usage, but the most
commonly used ones should be listed here. Users will rely on this for
quick reference when working with your software.

\textbf{Where to find more information}: Whether an email address, web
site or manual page, there should be an indication where the user can go
to find more information about the software.

\textbf{Printed to standard output} : So that it can be piped into
\texttt{less}, searched with \texttt{grep}, or compared to the previous
version with \texttt{diff}.

@@

\section{Version your releases}

Software evolves over time, with developers adding or removing features as need
dictates. Making official releases stamps a particular set of features with a
project-specific identifier so that version can be retrieved for later use. For
example, if a paper is published, the software should be released at the same
time so that the results can be reproduced. 
Most software has a version number composed of a decimal number that
increments as new versions are released.

There are many different ways
to construct and interpret the version, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
\textbf{Increment your version number every time you release your software to
other people}.

@@

\textbf{The version of your software should be easily available by 
supplying \texttt{-\/-version} or \texttt{-v} on the command line}. This command should print
the software name and version number, and it should
also be \textbf{included in all of the program's output}, particularly debugging
traces: if someone needs help, it's important that they be able to tell
whoever's helping them which version of the software they're using.

While new releases may make a program better in general,
they can simultaneously create work for someone
who integrated the old version into their own workflow a year or two ago,
and won't see any benefits from upgrading.
A program's authors should therefore \textbf{ensure that every released version
continues to be available.}
A number of mechanisms exist for
controlled release that range from as simple as adding an appropriate
commit message or tag to version control, to official releases alongside
code on Bitbucket or GitHub, to depositing into a
repository like apt, yum, homebrew, CPAN, etc. Choose the method that
best suits the number and expertise of users you anticipate.

@@

Despite these problems, software developers in research should
re-use existing software wherever possible and reasonable to avoid
``reinventing the wheel''. We fully support the reuse of software between
projects provided three additional guidelines are adhered to.

First, 
\textbf{make sure that you really need the auxiliary program}. If you are
executing GNU sort instead of figuring out how to sort lists in Python,
it may not be worth the pain of integration.

@@

Finally, to ensure support on as many different operating systems as
possible, \textbf{use native functions for starting other processes}, such as
Java's \texttt{Runtime.exec} call, Python's subprocess module, and Perl's system
command, and be sure to capture and report the output of the subprocess's standard error
to facilitate debugging.

@@

In all but the simplest of scripts, \textbf{we highly recommend using a build
automation utility to compile code and/or manage your software workflow}. The benefits of
using a build utility like Make, Rake, Maven, Ant or MS Build vastly outweighs the
drawbacks associated with the learning curve. Most importantly, a build system
lets someone else build your software the same way you would. Most build
utilities can also be configured to run tests during every build and
re-generate documentation, and can be hooked into continuous integration,
cutting releases, and deploying the software to other environments.

Build utilities rely on any dependencies being present in the environment for a
successful build, and so developers should
\textbf{document \emph{all} dependencies, preferably in a machine-readable form}.
Package managers like apt and yum are available on most *nix systems, and
application package managers exist for specific languages like Python (pip),
Java (Maven/Gradle), and Ruby (RubyGems). These package managers can be used
together with the build utility to ensure that dependencies are available at
compile/run time.
For example, it is common for Python projects to include a file called
\texttt{requirements.txt} that lists the names of required libraries,
along with version ranges:

@@

This file can be read by pip package manager, which can check that the
required software is available, and install it if it is not. 
Whatever is used,
we recommend that developers should \emph{always} install dependencies
using this mechanism, especially on their personal machines, so that
they're sure it works.

@@

Another reason for this rule is that users may want to try out a new
package before installing it system-wide on a cluster. Requiring root
privileges will frustrate such efforts, and thereby reduce uptake of the
package. Requiring that software be installed under its own user account
(e.g., that \texttt{packagename} be made a user, and all of the
package's software be installed in that ``user's'' space 
(see: Apache Tomcat))
is similarly
limiting, and makes side-by-side installation of multiple versions of
the package more difficult.

@@

For these reasons, users should be able to
\textbf{set the names and locations of input and output files as command-line parameters}.
This rule applies to reference data sets as well as the user's own
data: if a user wants to try a new gene identification algorithm using
a different set of genes as a training set, she should not have to
edit the software to do so.

A corollary to this rule is
\textbf{do not require users to navigate to a particular directory to do their work}.
``Where I have to be'' is just another hard-coded path.

@@

Useful parameters are those values that a user will need to modify to suit
their computer, dataset or application. 
The full list of useful parameters is software-specific and so cannot be
detailed here, but a short list includes providing
input and reference files and directories,
output files and directories,
filtering,
tuning,
random number generation seeds,
and
any alternatives that you've built-in,
e.g.,
compress results, use a different calculation, verbose output, etc.

Providing parameters on the
command line increases the flexibility and usability of the program. You
may have determined early in development that 0.58 is an optimal seed for your
original dataset, but that doesn't mean that is the best seed for every
case. Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
facilitating exploratory analysis and parameter sweeping.
\textbf{Choose reasonable defaults where they exist},
and \textbf{set no defaults at all when there aren't any reasonable ones}.

@@

Every package should come with a small test script for users to run
after installation. Its purpose is not only to check that the software
is working correctly (although that is extremely helpful), but rather to
ensure that it will work at all. This test script can also serve as a
working example of how to run the software.

In order to be useful, \textbf{this test script must be easy to find and run}. 
Many build systems will also run unit tests if provided them at compile time.
If the build system is not amenable to testing, a 
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a much better solution than documenting
test cases and requiring people to copy and paste them into the shell.

Equally, \textbf{the test script's output must be easy to interpret}. Screens
full of correlation coefficients do not qualify: instead, the script's
output should be simple to understand for non-experts,
such as one line per test, with the test's name
and its pass/fail status, followed by a single summary line saying how
many tests were run and how many passed or failed. If many or all tests
fail because of missing dependencies, that fact should be displayed
once, clearly, rather than once per test, so that users have a clear
idea of what they need to fix and how much work it's likely to take.

Studies have found that the ease with which people can start making
contributions is a strong predictor of whether they will or not~\cite{steinmacher2015}.
By making it simpler for outsiders to contribute,
a test suite of any kind also makes it more likely that they will, and software
with collaborators stands a better chance of surviving in the busy field of
scientific software.

@@

Many applications rely on randomized algorithms to
improve performance or runtimes. As a consequence, results can change
between runs, even when provided with the same data and parameters. By
its nature, this randomness renders strict reproducibility and therefore
debugging more difficult. If even the small test set (\#9) produces
different results, new users may not be able to tell whether the software is
working properly. When comparing
results between versions or after changing parameters, even small
differences can confuse or muddy the comparison. And especially when
producing results for publications, grants or diagnoses, any analysis
should be absolutely reproducible.

@@

Any short list of rules for developing robust software will
necessarily be incomplete.  A few of our more significant omissions
include:

>>>>

@@

Scientific software is typically developed and used by a single person,
usually a graduate student or postdoc~\cite{prins2015}.
It may produce the intended results in their hands,
but what happens when someone else wants to run it? Everyone with
a few years of experience feels a bit nervous when told to use
another person's code to analyze their data:
it will often be undocumented,
work in unexpected ways (if it works at all),
rely on nonexistent paths or resources,
be tuned for a single dataset,
or simply be an older version than was used in published papers.
The potential new user is then faced with two unpalatable options:
hack the existing code to make it work, or start over.

Being unable to
replicate results is so common that one publication refers to it as "a rite of
passage"~\cite{baker2016}. 
The root cause of this problem is that most research software
isn't \emph{robust}. The difference between running and being
robust is the difference between ``works for me on my machine'' and
``works for other people on a cluster I've never used''. The lack of
robustness also leads to many duplicated efforts,
which slows the pace of research~\cite{prabhu2011,lawlor2015}.
Bioinformatics software repositories \cite{ison2016,brazas2012} catalogue dozens to
hundreds of tools to perform the same task:
for example,
in 2016 the Bioinformatics Links Directory included 84 different multiple sequence aligners, 141 tools
to analyze transcript expression, and 182 pathway and interaction resources.
Some of these tools are legitimate efforts to improve the state-of-the-art, but
often they are difficult to install and run~\cite{stajich2002,Seemann2013}, and are effectively abandoned
after publication~\cite{nekrutenko2012}.

This problem is not unique to bioinformatics, or even to computing~\cite{baker2016}. 
Best practices in software engineering specifically aim to increase software
robustness. However, most bioinformaticians learn what they know about software development
on the job or otherwise informally~\cite{prins2015,atwood2015}.
Existing training programs and initiatives rarely have the time to cover software engineering
in depth, especially since the field is so broad and developing so rapidly~\cite{atwood2015,lawlor2015}.
In addition, making software robust is not directly rewarded
in science, and funding is difficult to come by~\cite{prins2015}. Some proposed
solutions to this problem include restructuring educational programs,
hiring dedicated software engineers~\cite{lawlor2015,sanders2008},
partnering with private sector or grassroots organizations~\cite{prins2015,ison2016},
or using specific technical tools like containerization or cloud
computing~\cite{afgan2016,howe2012}. Each of these requires time and, in some
cases, institutional change.

The good news is,
you don't need to be a professionally-trained programmer to write robust software. In fact,
some of the best, most reliable pieces of software in many scientific
communities are written by researchers~\cite{prabhu2011,sanders2008}
who have adopted strong software
engineering approaches, have high standards of reproducibility, use good testing
practices, and foster inter-institutional collaborations that grow into strong user
bases. In the bioinformatics community, Bioconductor and Galaxy follow this
path~\cite{gentleman2004,afgan2016}.

So what \emph{is} ``robust'' software?
In brief, it:

@@

These simple rules don't just ensure that the software will run for other people;
they are also necessary steps toward creating a reusable
library that can be shared and reused through a site like CPAN or CRAN.  They are
generic and can be applied to all languages, libraries, packages,
documentation styles, and operating systems for both closed-source and
open-source software.  
Whether the aim is as simple as sharing the
code with collaborators or as complex as using the software in a
production analysis environment, increasing the robustness of your
software decreases headaches all around.

\section{Have a README that briefly explains what the software does and what its dependencies are.}

The README is the first stop for most new users.
At a minimum, it needs to get a new user started and point them towards more
help, if they need it. 
Numerous guidelines exist on how to write good
READMEs\cite{Johnson1997,gnustandards};
their key common features are listed below.

\textbf{Explain what the software does:} At the beginning of the README,
explain what the software does in one or two sentences.
There's nothing more frustrating
than downloading and installing something only to
find out that it doesn't do what you thought it did.

@@

\textbf{Provide compilation/installation instructions:}
If the software needs to be compiled
or installed, list those instructions in the README. Also
mention if you recommend they use a pre-compiled binary instead
through a system such as pip, yum, or apt.

\textbf{List input and output files}: All possible input and output files
should be listed in this section. Do the files use a particular
standard, an extension of an existing format, or your own
format?  If they use a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
explicitly, listing all the required fields and acceptable values.
(You get bonus points if you include a script to convert between
a standard format and your file format). If there is no rigorous format
(which is common with log files), show at least a few lines from an
example file and explain what the sections mean.

Input files and their formats are included in most documentation, but
intermediate, auxiliary, and log files are often missing.
\emph{All} files should be listed in the README, even those considered self-explanatory. Log
files are often full of valuable information that can be
mined for the user's specific purpose. If your users might need to know,
``Does this program report the percentage of reads trimmed to remove
adapter sequences?'' they should be able to check the README and confidently
say, ``Yes, it is in the log file''.

\textbf{State attributions and licensing:} Attributions are how you credit
your main contributors; licenses are how others may use and
credit your work. These files should be available \emph{before}
the software is installed (or even downloaded).
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
Attributions can also contain a list of ``expert'' users
that can be contacted if new users have problems with the software.

@@

Usage information provides the first line of help for both first-time and
experienced users of command-line applications.
Ideally, usage is a terse, informative command-line help message that
guides the user in the correct use of the software. Terseness is
important: usage that extends for multiple screens is a nuisance, especially when
printed to standard error instead of standard output (where it can
easily be paged).  Long output is also much less likely to be read{\ldots}

Usage should provide all of the information necessary to run the
software. It is invoked either by running the software without
any arguments, running it with incorrect arguments, or by
explicitly choosing a help or usage option.
More standard command-line behaviours are detailed in \cite{Seemann2013}.

@@

\textbf{Most commonly used arguments, a description of each, and the
default values}: Not all arguments need to appear in the usage, but those most
commonly used should be listed. Users will rely on this for
quick reference when working with the software.

\textbf{Where to find more information}: Whether it's an email address, web
site, or manual page, there should be an indication where the user can go
to find out more about the software.

\textbf{Printed to standard output}: Usage should be displayed on standard output so that it can be piped into
\texttt{less}, searched with \texttt{grep}, or compared to the previous
version with \texttt{diff}.

@@

\section{Version your releases.}

Software evolves over time, with developers adding or removing features as need
dictates. Making official releases stamps a particular set of features with a
project-specific identifier so that version can be retrieved for later use. For
example, if a paper is published, the software should be released at the same
time so that the results can be reproduced.

Most software has a version number composed of a decimal number that
increments as new versions are released. 
There are many different ways
to construct and interpret this number, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
\textbf{Increment your version number every time you release your software to
other people}.

@@

\textbf{The version of your software should be easily available by 
supplying \texttt{-\/-version} or \texttt{-v} on the command line}. This command should print
the software name and version number, and it should
also be \textbf{included in all of the program's output}, particularly debugging
traces.  If someone needs help, it's important that they be able to tell
whoever's helping them which version of the software they're using.

While new releases may make a program better in general,
they can simultaneously create work for someone
who integrated the old version into their own workflow a year or two ago,
and won't see any benefits from upgrading.
A program's authors should therefore \textbf{ensure that old released versions
continue to be available.}
A number of mechanisms exist for
controlled release that range from as simple as adding an appropriate
commit message or tag to version control, to official releases alongside
code on Bitbucket or GitHub, to depositing into a
repository like apt, yum, homebrew, CPAN, etc. Choose the method that
best suits the number and expertise of users you anticipate.

@@

Despite these problems, software developers in research should
re-use existing software provided three guidelines are adhered to.
First, 
\textbf{make sure that you really need the auxiliary program}. If you are
executing GNU sort instead of figuring out how to sort lists in Python,
it may not be worth the pain of integration.

@@

Finally, to ensure support on as many different operating systems as
possible, \textbf{use native functions for starting other processes}, such as
Java's \texttt{Runtime.exec} call, Python's \texttt{subprocess} module, and Perl's \texttt{system}
command, and be sure to capture and report the output of the subprocess's standard error
to facilitate debugging.

@@

In all but the simplest of scripts, \textbf{use a build
automation utility to compile code and/or manage your software workflow}. The benefits of
using tools like Make, Rake, Maven, Ant or MS Build vastly outweigh the
their learning curves. Most importantly, a build system
lets someone else build your software the same way you would. Most build
utilities can also be configured to run tests during every build and
re-generate documentation, and can be hooked into continuous integration,
cutting releases, and deploying the software to other environments.

Build utilities rely on any dependencies being present in the environment for a
successful build, and so developers should
\textbf{document \emph{all} dependencies, preferably in a machine-readable form}.
Package managers like apt and yum are available on most Unix-like systems, and
application package managers exist for specific languages like Python (pip),
Java (Maven/Gradle), and Ruby (RubyGems). These package managers can be used
together with the build utility to ensure that dependencies are available at
compile/run time.
For example, it is common for Python projects to include a file called
\texttt{requirements.txt} that lists the names of required libraries,
along with version ranges:

@@

This file can be read by the pip package manager, which can check that the
required software is available and install it if it is not. 
Whatever is used,
we recommend that developers should \emph{always} install dependencies
using their dependency description, especially on their personal machines, so that
they're sure it works.

@@

Another reason for this rule is that users may want to try out a new
package before installing it system-wide on a cluster. Requiring root
privileges will frustrate such efforts, and thereby reduce uptake of the
package. Requiring, as Apache Tomcat does, that software be installed
under its own user account---i.e.,
that \texttt{packagename} be made a user, and all of the
package's software be installed in that pseudo-user's space---is similarly limiting,
and makes side-by-side installation of multiple versions of
the package more difficult.

@@

For these reasons, users should be able to
\textbf{set the names and locations of input and output files as command-line parameters}.
This rule applies to reference data sets as well as the user's own
data: if a user wants to try a new gene identification algorithm using
a different set of genes as a training set, she should not have to
edit the software to do so.
A corollary to this rule is
\textbf{do not require users to navigate to a particular directory to do their work},
since ``where I have to be'' is just another hard-coded path.

@@

Useful parameters are those values that a user will need to modify to suit
their computer, dataset or application. 
The full list of useful parameters is software-specific and so cannot be
detailed here, but a short list includes providing
input and reference files and directories,
output files and directories,
filtering,
tuning,
random number generation seeds,
and
alternatives such as compressing results,
use a variant algorithm,
or verbose output.

Providing parameters on the
command line increases the flexibility and usability of the program. You
may have determined early in development that 0.58 is an optimal cluster threshold for your
original dataset, but that doesn't mean it is best for every
case. Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
as it facilitates exploratory analysis and parameter sweeping.
\textbf{Choose reasonable defaults where they exist},
and \textbf{set no defaults at all when there aren't any reasonable ones}.

@@

Every package should come with a small test script for users to run
after installation. Its purpose is not only to check that the software
is working correctly (although that is extremely helpful), but also to
ensure that it works at all. This test script can also serve as a
working example of how to run the software.

In order to be useful, \textbf{make the test script easy to find and run}. 
Many build systems will also run unit tests if provided them at compile time.
If the build system is not amenable to testing, a 
single file in the project's root directory named \texttt{runtests.sh}
or something equally obvious is a much better solution than documenting
test cases and requiring people to copy and paste them.

Equally, \textbf{make the test script's output easy to interpret}. Screens
full of correlation coefficients do not qualify: instead, the script's
output should be simple to understand for non-experts,
such as one line per test, with the test's name
and its pass/fail status, followed by a single summary line saying how
many tests were run and how many passed or failed. If many or all tests
fail because of missing dependencies, that fact should be displayed
once, clearly, rather than once per test, so that users have a clear
idea of what they need to fix and how much work it's likely to take.

Research has shown that the ease with which people can start making
contributions is a strong predictor of whether they will or not~\cite{steinmacher2015}.
By making it simpler for outsiders to contribute,
a test suite of any kind also makes it more likely that they will, and software
with collaborators stands a better chance of surviving in the busy field of
scientific software.

@@

Many applications rely on randomized algorithms to
improve performance or runtimes. As a consequence, results can change
between runs, even when provided with the same data and parameters. By
its nature, this randomness renders strict reproducibility and therefore
debugging more difficult. If even the small test set (\#9) produces
different results for each run, new users may not be able to tell whether the software is
working properly. When comparing
results between versions or after changing parameters, even small
differences can confuse or muddy the comparison. And especially when
producing results for publications, grants or diagnoses, any analysis
should be absolutely reproducible.

@@

Any short list of rules for developing robust software will
necessarily be incomplete.   Our more significant omissions
include:
